---
title: "Move points to nearest raster cell"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(rSDM)
```

`rSDM` includes a handy function to move points to the nearest raster cell with data (i.e. not NA).

## Example data

Let's generate example point coordinates:

```{r}
locs <- data.frame(lon = c(1, 2, 1, 2, 2.2), lat = c(1.2, 1, 2.3, 3, 2))
locs.sf <- locs2sf(locs)
locs.sf
```

Now let's generate an example raster:

```{r}
library(terra)
ras <- rast(nrows = 2, ncols = 2, xmin = 0.5, xmax = 3.5, ymin = 0.5, ymax = 3.5,
 resolution = 1, vals = c(NA, 1, 1, NA, NA, 1, NA, 1, 1))
```

As we can see in the map below, some points fall outside raster cells with data:

```{r}
occmap(locs.sf, ras, pcol = "black", psize = 3)
```

## Move point coordinates to nearest raster cell

```{r}
moved <- points2nearestcell(locs.sf, ras)
```

Note you can choose the format of the output map, e.g. now using ggplot (interactive leaflet maps are available too):

```{r}
points2nearestcell(locs.sf, ras, map = "ggplot")
```

The function returns a spatial object where the coordinates of the points falling outside the raster have been moved:

```{r}
moved
```

Let's compare the original and new coordinates:

```{r}
sf::st_coordinates(locs.sf)
sf::st_coordinates(moved)
```

If you don't want to change coordinates but only check which would the nearest raster cells for each point, use `move = FALSE`.

## Move points to nearest raster cell within a given distance

In case you want to move points **only** if the nearest raster cell is within a given distance, you could use the `distance` argument to set a threshold:

```{r}
moved <- points2nearestcell(locs.sf, ras, distance = 100000)
```


