---
title: "Map occurrences (point coordinates)"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(rSDM)
```

Let's create example point coordinates and convert them to a spatial (sf) object:

```{r}
locs <- data.frame(lon = c(-5.8, -5.5, -5.9), lat = c(35.7, 36.2, 36.5))
locs <- locs2sf(locs, crs = 4326)
locs
```

Note that SpatVector objects (from terra) are also accepted, and can be created using `rSDM::locs2vect()`.

## Default map with occmap

Unless a raster `ras` is provided, `occmap` will automatically download a background map from the internet. 

```{r fig.keep='last'}
occmap(locs, psize = 6)
```

The background map can be easily changed using the `bg` argument:

```{r fig.keep='last'}
occmap(locs, psize = 6, bg = "CartoDB.Positron")
```



## Interactive (leaflet) map

```{r}
occmap(locs, psize = 6, type = "leaflet")
occmap(locs, psize = 6, type = "leaflet", bg = "CartoDB.Positron")
```


## ggplot map

```{r}
occmap(locs, psize = 6, type = "ggplot")
```


## Adding points to a previous map

Let's create some new coordinates (e.g. for a different species):

```{r}
new.locs <- data.frame(lon = c(-5.8, -5.4), lat = c(36.2, 36.5))
new.locs.sf <- locs2sf(new.locs, crs = 4326)
```

### Using 'base' maps

Note the `add = TRUE` argument:

```{r fig.keep='last'}
map <- occmap(locs, psize = 6)
occmap(new.locs.sf, add = TRUE, psize = 6, pcol = "blue")
```


### Adding points to a previous map (leaflet)

Note the previous map has to be provided (using `prev.map` argument):

```{r}
map <- occmap(locs, psize = 6, type = "leaflet")
occmap(new.locs.sf, prev.map = map, psize = 6, pcol = "blue")
```


### Adding points to a previous map (ggplot)

Note the previous map has to be provided (using `prev.map` argument):

```{r}
map <- occmap(locs, psize = 6, type = "ggplot")
occmap(new.locs.sf, prev.map = map, psize = 6, pcol = "blue")
```


